### 2017-04-30 paper/jfp-submission/interactive-short.tex
Wes pass over the the entire document. Minor editing changes only.

\paragraph{Jump Compression}
%
A \emph{jump compressed} trace is one whose edges are limited to forward
or backward jumps.
% (The sizes of these traces is equal to the jump
% metric below.)
%
In our experience, jump compression abstracts many details of the
computation that are often uninteresting or irrelevant to the
explanation.
%
In particular, jump compressed traces hide low-level operations and
summarize function calls as call-return pairs, see
Figure~\ref{fig:fac-jump} for a variant of @fac@ that implements the
subtraction as a function call instead of a primitive.
%
\begin{figure}[t]
\centering
\includegraphics[height=2in]{fac-minus.png}
\caption{Jump-compressed trace of \texttt{fac 1} with subtraction
  implemented as a function call.}
\label{fig:fac-jump}
\end{figure}
%
Once users have identified interesting call-return pairs, they can
step into those calls and proceed with more fine-grained steps.
%
% Expanding the initial visualization to a jump-compressed trace is often
% an ideal first step;
% % expanding a jump-compressed traces ideal for the initial state of the
% they provide a concise overview of the execution and allow the user to
% step directly into an interesting function call, at which point the can
% proceed with more fine-grained steps.
%
Note that jump compressed traces are not quite the same as
stack-traces as they show \emph{all} function calls, including
those that returned successfully.
%
% For example, the display on the bottom-left in Figure~\ref{fig:factorial} shows
% a jump-compressed trace, of size 4. The full trace for
% the same example (on the right of the Figure~\ref{fig:factorial}) has size 19.

>>>>

\paragraph{Jump Compression}
%
A \emph{jump compressed} trace is one whose edges are limited to forward
or backward jumps.
% (The sizes of these traces is equal to the jump
% metric below.)
%
In our experience, jump compression abstracts many details of the
computation that are often uninteresting or irrelevant to the
explanation.
%
In particular, jump compressed traces hide low-level operations and
summarize function calls as call-return pairs, see
Figure~\ref{fig:fac-jump} for a variant of @fac@ that implements the
subtraction as a function call instead of a primitive.
%
\begin{figure}[t]
\centering
\includegraphics[height=2in]{fac-minus.png}
\caption{Jump-compressed trace of \texttt{fac 1} with subtraction
  implemented as a function call.}
\label{fig:fac-jump}
\end{figure}
%
Once users have identified interesting call-return pairs, they can
step into those calls and proceed with more fine-grained steps.
%
% Expanding the initial visualization to a jump-compressed trace is often
% an ideal first step;
% % expanding a jump-compressed traces ideal for the initial state of the
% they provide a concise overview of the execution and allow the user to
% step directly into an interesting function call, at which point the can
% proceed with more fine-grained steps.
%
Note that jump compressed traces are not quite the same as
stack traces as they show \emph{all} function calls, including
those that returned successfully.
%
% For example, the display on the bottom-left in Figure~\ref{fig:factorial} shows
% a jump-compressed trace, of size 4. The full trace for
% the same example (on the right of the Figure~\ref{fig:factorial}) has size 19.

### 2017-04-30 paper/jfp-submission/introduction.tex
mention localization in abs/intro
Wes pass over the the entire document. Minor editing changes only.

We introduce a notion of \emph{jump-compressed} traces to abstract away
the irrelevant details of a computation.
%
A jump-compressed trace includes only function
calls and returns, for example the trace in the bottom-left of
Figure~\ref{fig:factorial} is jump-compressed.
%
Jump-compressed traces are similar to stack traces, both show a
sequence of function calls that lead to a crash, but the jump-compressed
trace also shows the return values of successful calls, which can be
useful in understanding why a particular path was taken.

\paragraph{3. Evaluating Witnesses}
%
Of course, the problem of finding witnesses is
undecidable in general. In fact, due to the necessarily
conservative nature of static typing, there
may not even exist any witnesses for a given
ill-typed program.
%
Thus, our approach is a heuristic that is only useful
if it can find \emph{compact} witnesses for
\emph{real-world} programs.
%
Our third contribution is an extensive evaluation of our approach
on two different sets of ill-typed programs obtained by instrumenting
compilers used in beginner's classes~(\S~\ref{sec:evaluation}).
%
The first is the \uwbench\ data set~\cite{Lerner2007-dt}
comprising \uwsize\ ill-typed programs.
%
The second is a new \ucsdbench\ data set, comprising \ucsdsize\
ill-typed programs.
%
We show that for both data sets, our technique is able to generate
witnesses for around 85\% of the programs, in under a second in the
vast majority of cases.
%
Furthermore, we show that a simple interactive strategy yields
compact counterexample traces with at most 5 steps for 60\%
of the programs, and at most 10 steps for over 80\% of the programs.

>>>>

We introduce a notion of \emph{jump-compressed} traces to abstract away
the irrelevant details of a computation.
%
A jump-compressed trace includes only function
calls and returns. For example, the trace in the bottom-left of
Figure~\ref{fig:factorial} is jump-compressed.
%
Jump-compressed traces are similar to stack traces in that both show a
sequence of function calls that lead to a crash. However, jump-compressed
traces also show the return values of successful calls, which can be
useful in understanding why a particular path was taken.

\paragraph{3. Evaluating Witnesses}
%
Of course, the problem of finding witnesses is
undecidable in general. In fact, due to the necessarily
conservative nature of static typing, there
may not even exist any witnesses for a given
ill-typed program.
%
Thus, our approach is a heuristic that is only useful
if it can find \emph{compact} witnesses for
\emph{real-world} programs.
%
Our third contribution is an extensive evaluation of our approach
on two different sets of ill-typed programs obtained by instrumenting
compilers used in beginner's classes~(\S~\ref{sec:evaluation}).
%
The first is the \uwbench\ data set~\cite{Lerner2007-dt}
comprising \uwsize\ ill-typed programs.
%
The second is a new \ucsdbench\ data set, comprising \ucsdsize\
ill-typed programs.
%
We show that for both data sets, our technique is able to generate
witnesses for around 85\% of the programs, in under a second in the
vast majority of cases.
%
Furthermore, we show that a simple interactive strategy yields
compact counterexample traces with at most 5 steps for 60\%
of the programs, and at most 10 steps for over 80\% of the programs.
%
We can even use witnesses to \emph{localize} type errors with a simple
heuristic that treats the values in a ``stuck'' term as \emph{sources}
of typing constraints and the term itself as a \emph{sink},
achieving around 70\% accuracy in locating the source of the error.

### 2017-04-30 paper/jfp-submission/safe.tex
misc tweaks
tweak
Wes pass over the the entire document. Minor editing changes only.

\subsubsection{How safe are the ``safe'' programs?}
\label{sec:how-safe}

@@

\begin{figure}[t]
%\centering
\includegraphics[width=0.8\linewidth]{distrib_ext.png}
\vspace{-0.75cm}
\caption{Results of our investigation into programs where \toolname
  did not produce a witness. A ``*'' denotes that the percentage is an
  estimate based on a random sampling of 50 programs.}
\label{fig:no-witness}
\end{figure}

@@

Extending \toolname with support for symbolic execution would alleviate
this issue, as we could then begin symbolically executing the program
until we learn how to instantiate |n| and |m|.
%
Alternatively, we could \emph{speculatively} instantiate both |n| and
|m| with some arbitrary type, and proceed with execution until we
discover a type error.
%
This speculative instantiation is of course unsound; we would have to
take care to avoid reporting frivolous type errors that were caused by
such instantiations.
%
We would need to track which holes were instantiated speculatively in
order to distinguish type errors that would have happened regardless, as
in @fac@, from type errors that were caused by our instantiation.

Further, suppose that our speculative instantiation induces a frivolous
type error.
%
For example, suppose we are given
%
\begin{code}
  let bad x y =
    if x < y then
      x *. y
    else
      0.0
\end{code}
%
and choose to (speculatively) instantiate @x@ and @y@ as @int@s and proceed
down the ``true'' branch.
%
We will quickly discover this was the wrong choice, as they are immediately
narrowed to @float@s.
%
We must now backtrack and try a different instantiation, but we no
longer need to choose one at random.
%
Since our instantiation was speculative, we can still view @x@ and @y@
\emph{morally} as holes, and thus treat the @*.@ operator as a normal
narrowing point with two holes.
%
This tells us that the \emph{correct} instantiation was in fact @float@,
and we can then proceed as normal from the backtracking point with a
concrete choice of @float@s.
%
Thus, it appears that speculative instantiation of holes may be a
useful, lightweight alternative to symbolic execution for our purposes.

@@

Thus, it is not particularly satisfying that \toolname fails to produce
a witness here; a possible solution could be to track the types of the
inputs, and demonstrate to the user how they change between recursive
calls.
%
This would require maintaining a typing environment of variables in
addition to the environments we maintain for holes.
%
We would have to modify the rule $\reappgood$ from
Figure~\ref{fig:operational} to additionally $\forcesym$ the function's
type against the concrete inputs.
%
However, we would want to ensure that this $\forcesym$ cannot fail ---
it is preferable to report a stuck term as that provides a fuller view
of the error.
%
Rather, we would note which evaluation steps induced incompatible
type refinements, and if a traditional witness cannot be found, we could
then report a trace expanded to show precisely these steps.
%
This seems like a modest extension to our semantics, and would be
interesting to explore further.

>>>>

\subsection{How safe are the ``safe'' programs?}
\label{sec:how-safe}

@@

\begin{figure}[t]
%\centering
\includegraphics[width=0.7\linewidth]{distrib_ext.png}
\vspace{-0.75cm}
\caption{Results of our investigation into programs where \toolname
  did not produce a witness. A ``*'' denotes that the percentage is an
  estimate based on a random sampling of 50 programs.}
\label{fig:no-witness}
\end{figure}

@@

Extending \toolname with support for symbolic execution would alleviate
this issue, as we could then begin symbolically executing the program
until we learn how to instantiate |n| and |m|.
%
Alternatively, we could \emph{speculatively} instantiate both |n| and
|m| with some arbitrary type, and proceed with execution until we
discover a type error.
%
This speculative instantiation is, of course, unsound; we would have to
take care to avoid reporting frivolous type errors that were caused by
such instantiations.
%
We would need to track which holes were instantiated speculatively 
to distinguish type errors that would have happened regardless, as
in @fac@, from type errors that were caused by our instantiation.

Further, suppose that our speculative instantiation induces a frivolous
type error.
%
For example, suppose we are given
%
\begin{code}
  let bad x y =
    if x < y then
      x *. y
    else
      0.0
\end{code}
%
and choose to (speculatively) instantiate @x@ and @y@ as @int@s and proceed
down the ``true'' branch.
%
We will quickly discover this was the wrong choice, as they are immediately
narrowed to @float@s.
%
We must now backtrack and try a different instantiation, but we no
longer need to choose one at random.
%
Since our instantiation was speculative, and @x@ and @y@
% WRW thinks "morally" is too informal for this venue. 
were \emph{originally} hole, we can treat the @*.@ operator as a normal
narrowing point with two holes.
%
This tells us that the \emph{correct} instantiation was in fact @float@,
and we can then proceed as normal from the backtracking point with a
concrete choice of @float@s.
%
Thus, it appears that speculative instantiation of holes may be a
useful, lightweight alternative to symbolic execution for our purposes.

@@

Thus, it is not particularly satisfying that \toolname fails to produce
a witness here; a possible solution could be to track the types of the
inputs, and demonstrate to the user how they change between recursive
calls.
%
This would require maintaining a typing environment of variables in
addition to the environments we maintain for holes.
%
We would have to modify the rule $\reappgood$ from
Figure~\ref{fig:operational} to additionally $\forcesym$ the function's
type against the concrete inputs.
%
However, we would want to ensure that this $\forcesym$ cannot fail ---
it is preferable to report a stuck term as that provides a fuller view
of the error.
%
Rather, we would note which evaluation steps induced incompatible
type refinements, and if a traditional witness cannot be found, we could
then report a trace expanded to show precisely these steps.
%
This represents only a modest extension to our semantics, and would be
interesting to explore further.
